Reproducible Research: Peer Assessment 1
=======================================

## Loading and preprocessing the data

Unzip the file and load the csv into R.
```{r load, echo=TRUE}
unzip("activity.zip")
activity <- read.csv("activity.csv")
head(activity)
```

Use dplyr for this analysis.
```{r dplyr, echo=TRUE}
library(dplyr)
```

## What is mean total number of steps taken per day?

A quick look at the daily number steps distribution.
```{r histogram, echo=TRUE, fig.height=4}
library(ggplot2)
steps_by_day <- activity %>% group_by(date) %>% summarize(steps=sum(steps))
qplot(steps, data=steps_by_day, geom="histogram",binwidth=1000)
m <- mean(steps_by_day$steps, na.rm=TRUE)
md <- median(steps_by_day$steps, na.rm=TRUE)
```

The average number of steps per day is `r format(m)` and the median is `r md`.

## What is the average daily activity pattern?

```{r intervals, echo=TRUE, fig.height=4}
by_interval <- activity %>% group_by(interval) %>% summarize(steps=mean(steps, na.rm=TRUE))
with( by_interval, plot(interval, steps, type="l"))
```

There seems to be a spike in activity in the morning.

The maximum number of steps over a 5 minutes interval is obtained in interval beginning at:
```{r argmax, echo=TRUE}
by_interval$interval[which.max(by_interval$steps)]
```

## Imputing missing values

There are a high number of intervals with missing values:
```{r missing, echo=TRUE}
sum(is.na(activity$steps))
```

```{r missing analysis, echo=TRUE}
missing <- activity %>% filter(is.na(steps)) %>% select(date, interval)
summary(missing, maxsum=10)
```

The missing values are concentrated on 9 days for wi=hich there are no measurements at all.

In this case a simple imputing strategy could be to use the median value over that interval.
```{r imputing, echo=TRUE}
median_by_interval <- activity %>% group_by(interval) %>% summarise(steps=median(steps, na.rm=TRUE))
impute <- function(steps, interval) {
    if(is.na(steps)) {
        median_by_interval$steps[median_by_interval$interval==interval]
    } else {
        steps
    }
}
imputed <- activity
imputed$steps <- mapply(impute,imputed$steps, imputed$interval)
```

```{r evaluation, echo=TRUE, fig.height=4}
imputed_by_day <- imputed %>% group_by(date) %>% summarize(steps=sum(steps))
qplot(steps, data=imputed_by_day, geom="histogram",binwidth=1000)
m <- mean(imputed_by_day$steps)
md <- median(imputed_by_day$steps)
```

After imputation the average number of steps per day is `r format(m)` and the median is `r md`.

Using the interval median for imputation seems to have reduced both the average and the median number of steps per day. 

## Are there differences in activity patterns between weekdays and weekends?
```{r weekdays, echo=TRUE}
imputed <- imputed %>%
    mutate(wkday = weekdays(strptime(imputed$date,format="%Y-%m-%d"))) %>%
    mutate(wkend = ifelse(wkday=="Sunday"|wkday=="Saturday", "weekend", "weekday"))
by_interval <- imputed %>% group_by(interval, wkend) %>% summarize(steps=mean(steps))
qplot(interval,steps,data=by_interval,facets=wkend~., geom="line")
```
